EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-05-19 draft.tex
Spell check

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration is generated
by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deploy the controller, setup and redis 
containers we used the file in Listing ~\ref{code:compose}.
Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

We deployed two ECS configurations, one that included all containers in a single task,
and another one, where we assigned each worker to a different task and the controller, 
redis, and setup in another, for a total of nine concurrent Fargate tasks.

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective ben chmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.

@@

We have deployed the EvoSwarm application locally in the same hardware and
software described in Section~\ref{setup:local-cloud}, additionally the BOBB
experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
in Python, the plots were produced with version 2.3.2.

@@

\begin{table}[h!tbp]
  \small
  \caption{Speedup achieved with incrementing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{lllllll}
            &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
  Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
  2         & 1w            & 2w            & {\ul 1.94}            & 6E-16           & 1.84             & 2E-16           \\
            & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
            & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
  3         & 1w            & 2w            & {\ul 1.93}            & 2E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
            & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
  5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
            & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
            & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
  10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.09}      & 1E-03           & 1.78             & 2E-16           \\
  20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
            & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.04 }     & 1E-02           & 1.80             & 2E-16           \\
  40        & 1w            & 2w            & \textbf{2.00}   & 2E-16           & 1.95             & 2E-16           \\
            & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.07}            & 3E-02           & \textbf{1.98}    & 2E-16          
  \end{tabular}
  \end{table}
%
Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
one worker. When using eight workers, we achieved the maximum speed up for five
and ten populations, but interestingly, it was in dimension 40 and dimension 2 
(these values are shown in bold in the table).

@@

In order to run this experiment, we just needed to edit a configuration file and
start it up. The plots presented in the results are generated by the BBOB
framework from the JOSN log file for each run. There was no additional setup
needed, proving how this framework is able to not only offer an reproducible way
to run experiments, but also how easily it interfaces with other common tools in
the EC ecosystem.

>>>>

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration of a task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration is generated
by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
the configuration on-the-fly. For example to deploy the controller, setup and Redis 
containers we used the file in Listing ~\ref{code:compose}.
Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
that reference the container images we have published in Docker hub
\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
variables used by code running inside containers.

We deployed two ECS configurations, one that included all containers in a single task,
and another one, where we assigned each worker to a different task and the controller, 
Redis, and setup in another, for a total of nine concurrent Fargate tasks.

@@

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.

@@

We have deployed the EvoSwarm application locally in the same hardware and
software described in Section~\ref{setup:local-cloud}, additionally the BBOB
experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
in Python, the plots were produced with version 2.3.2.

@@

\begin{table}[h!tbp]
  \small
  \caption{Speedup achieved with increasing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{lllllll}
            &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
  Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
  2         & 1w            & 2w            & {\ul 1.94}            & 6E-16           & 1.84             & 2E-16           \\
            & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
            & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
  3         & 1w            & 2w            & {\ul 1.93}            & 2E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
            & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
  5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
            & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
            & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
  10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.09}      & 1E-03           & 1.78             & 2E-16           \\
  20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
            & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.04 }     & 1E-02           & 1.80             & 2E-16           \\
  40        & 1w            & 2w            & \textbf{2.00}   & 2E-16           & 1.95             & 2E-16           \\
            & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.07}            & 3E-02           & \textbf{1.98}    & 2E-16          
  \end{tabular}
  \end{table}
%
Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
one worker. When using eight workers, we achieved the maximum speed up for five
and ten populations, but interestingly, it was in dimension 40 and dimension 2 
(these values are shown in bold in the table).

@@

In order to run this experiment, we just needed to edit a configuration file and
start it up. The plots presented in the results are generated by the BBOB
framework from the JSON log file for each run. There was no additional setup
needed, proving how this framework is able to not only offer an reproducible way
to run experiments, but also how easily it interfaces with other common tools in
the EC ecosystem.
